---
title: "Premier League Match Prediction - Poisson Model of Goals Scored"
author: "Craig McGowan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(pander)
library(zoo)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "../")
```

```{r helper functions}

# Calculate recent form for a particular stat
recent_form <- function(var, num_weeks, type) {

  if (type == "sum") {
    lag(zoo::rollsum(var, num_weeks, na.pad = TRUE, align = "right"), 1)
  }
  if (type == "mean") {
    lag(zoo::rollmean(var, num_weeks, na.pad = TRUE, align = "right"), 1)
  }
}

# Fit Poisson model and cross-validate
poisson_model <- function(x, features) { 
 
  # Remove any missing values
  x <- x %>%
    select(season, date, hg, ag, res, one_of(features)) %>%
    na.omit()
  
  season_results <- list()
  for (this_season in unique(x$season)[1]) {
    train <- filter(x, season != this_season)
    test <- filter(x, season == this_season)
 
    home_mod <- glm(formula(paste("hg ~ ", paste0(features, collapse = " + "))),
                    family = "poisson", data = train)
    away_mod <- glm(formula(paste("ag ~ ", paste0(features, collapse = " + "))),
                    family = "poisson", data = train)
    
    train$home_exp_goals <- predict(home_mod, newdata = train, type = "response")
    train$away_exp_goals <- predict(away_mod, newdata = train, type = "response")
    
    test$home_exp_goals <- predict(home_mod, newdata = test, type = "response")
    test$away_exp_goals <- predict(away_mod, newdata = test, type = "response")
    
    season_results[[this_season]] <- train %>%
      rowwise() %>%
      mutate(pred_result = pred_result(home_exp_goals, away_exp_goals)) %>%
      ungroup() %>%
      summarize(train_result = mean(pred_result == res)) %>%
      cbind(test %>%
        rowwise() %>%
        mutate(pred_result = pred_result(home_exp_goals, away_exp_goals)) %>%
        ungroup() %>%
        summarize(test_result = mean(pred_result == res)))
  }
  
  # Output average of cross-validated results
  bind_rows(season_results) %>%
    summarize_all(mean) %>%
    mutate(features = paste0(features, collapse = ", "))
}

# Create matrix of probabilities of goals and determine predicted result
pred_result <- function(team_exp_goals, opp_exp_goals) {
  pred_matrix <- matrix(nrow = 6, ncol = 6)
  
  for (i in 1:6) {
    for (j in 1:6) {
      pred_matrix[i, j] = dpois(i-1, team_exp_goals) * 
        dpois(j-1, opp_exp_goals)
    }
  }
  prob_draw <- sum(diag(pred_matrix))
  
  prob_home <- sum(pred_matrix[lower.tri(pred_matrix)])
  
  prob_away <- sum(pred_matrix[upper.tri(pred_matrix)])
  
  case_when(
    prob_draw > prob_home & prob_draw > prob_away ~ "D",
    prob_home > prob_draw & prob_home > prob_away ~ "H",
    prob_away > prob_home & prob_away > prob_draw ~ "A",
    TRUE ~ NA_character_
  )
}
```


```{r data setup}
load("Data/prem.Rdata")
load("Data/prem_adjusted.Rdata")

prem_league <- prem_league %>%
  left_join(select(prem_adjust_goals, -season, -hg, -ag, -res),
            by = c("date", "home", "away"))

# Dataset of team characteristics
team_char <- prem_league %>%
  select(season, date, team = home, goal_for = hg, 
         goal_against = ag, shot_for = hs, shot_against = as,
         shot_target_for = hst, shot_target_against = ast,
         ags = home_ags, aga = home_aga) %>%
  # Add location-specific match number
  group_by(season, team) %>%
  arrange(date) %>%
  mutate(location = "home",
         loc_match_number = row_number()) %>%
  ungroup() %>%
  bind_rows(prem_league %>%
              select(season, date, team = away, goal_for = ag, 
                     goal_against = hg, shot_for = as, shot_against = hs,
                     shot_target_for = ast, shot_target_against = hst,
                     ags = away_ags, aga = away_aga) %>%
              # Add location-specific match number
              group_by(season, team) %>%
              arrange(date) %>%
              mutate(location = "away",
                     loc_match_number = row_number()) %>%
              ungroup()) %>%
  # Add match week
  group_by(season, team) %>%
  arrange(date) %>%
  mutate(match_week = row_number()) %>%
  ungroup() %>%
  # Calculate season summary stats
  group_by(season, team, location) %>%
  mutate(ssn_team_avg_gf = mean(goal_for),
         ssn_team_avg_ga = mean(goal_against)) %>%
  ungroup() %>%
  # Calculate long-term rolling averages of goals and shots
  # 20 match rolling mean across seasons, location specific
  group_by(team, location) %>%
  mutate(roll20_gf = rollmean(lag(goal_for, 1), 
                                   20, align = "right", fill = NA),
         roll20_ga = rollmean(lag(goal_against, 1),
                                   20, align = "right", fill = NA)) %>%
  ungroup() %>%
  # 20 match rolling mean of shots allowed across seasons, location specific
  group_by(team, location) %>%
  mutate(roll20_sf = rollmean(lag(shot_for, 1), 
                                   20, align = "right", fill = NA),
         roll20_sa = rollmean(lag(shot_against, 1),
                                   20, align = "right", fill = NA)) %>%
  ungroup() %>%
  # Recent 5 match in-season form  - goals
  group_by(season, team, location) %>%
  mutate(recent5_gf = lag(zoo::rollsum(goal_for, 5, 
                                            na.pad = TRUE, align = "right"), 1),
         recent5_ga = lag(zoo::rollsum(goal_against, 5, 
                                            na.pad = TRUE, align = "right"), 1)) %>%
  ungroup() %>%
  # Recent 5 match in-season form  - adjusted goals
  group_by(season, team) %>%
  mutate(recent5_ags = lag(zoo::rollsum(ags, 5,
                                        na.pad = TRUE, align = "right"), 1),
         recent5_aga = lag(zoo::rollsum(aga, 5,
                                        na.pad = TRUE, align = "right"), 1)) %>%
  ungroup() %>%
  # Recent 5 match in-season form  - shots
  group_by(season, team, location) %>%
  mutate(recent5_sf = lag(zoo::rollsum(shot_for, 5, 
                                            na.pad = TRUE, align = "right"), 1),
         recent5_sa = lag(zoo::rollsum(shot_against, 5, 
                                            na.pad = TRUE, align = "right"), 1)) %>%
  ungroup()


# Bind team characteristics onto match outcomes
model_df <- prem_adjust_goals %>%
  left_join(team_char %>%
              filter(location == "home") %>%
              select(date, home = team, home_loc_match_number = loc_match_number,
                     home_match_number = match_week, home_roll20_gf = roll20_gf,
                     home_roll20_ga = roll20_ga, home_roll20_sf = roll20_sf,
                     home_roll20_sa = roll20_sa, home_recent5_gf = recent5_gf,
                     home_recent5_ga = recent5_ga, home_recent5_sf = recent5_sf,
                     home_recent5_sa = recent5_sa, home_recent5_aga = recent5_aga,
                     home_recent5_ags = recent5_ags),
            by = c("date", "home")) %>%
  left_join(team_char %>%
              filter(location == "away") %>%
              select(date, away = team, away_loc_match_number = loc_match_number,
                     away_match_number = match_week, away_roll20_gf = roll20_gf,
                     away_roll20_ga = roll20_ga, away_roll20_sf = roll20_sf,
                     away_roll20_sa = roll20_sa, away_recent5_gf = recent5_gf,
                     away_recent5_ga = recent5_ga, away_recent5_sf = recent5_sf,
                     away_recent5_sa = recent5_sa, away_recent5_aga = recent5_aga,
                     away_recent5_ags = recent5_ags),
            by = c("date", "away"))
  
```

Outcome - goals scored
Features - home/away, measure of opponent defensive strength, measure of team strength,
  recent form (goals, shots),


# Model 1
Home and away offensive and defensive ratings
```{r mod 1}
poisson_model(x = model_df %>% filter(season != "9394"), 
              features = c("h_off", "h_def", "a_off", "a_def")) %>%
  pander()
```

# Model 2
5 match rolling mean goals scored (location specific), 5 match rolling mean opponent goals allowed (location specific)
```{r mod 2}
poisson_model(x = model_df ,
              features = c("home_recent5_gf", "home_recent5_ga",
                           "away_recent5_gf", "away_recent5_ga")) %>%
  pander()
```

# Model 3
5 match rolling mean adjusted goals scored and adjusted goals allowed.
```{r mod 3}
poisson_model(x = model_df ,
              features = c("home_recent5_ags", "home_recent5_aga",
                           "away_recent5_ags", "away_recent5_aga")) %>%
  pander()
```