---
title: "Premier League Model Match Prediction"
author: "Craig McGowan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(pander)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "../")
```

```{r functions}
# Recent form for a particular stat
recent_form <- function(var, num_weeks) {

  lag(zoo::rollsum(var, num_weeks, na.pad = TRUE, align = "right"), 1)
 
}

single_recent_model <- function(df) {
  
  # Combine calculated form with observed results
  model_df <- prem_combined %>%
    select(date, season, home, away, res) %>%
    # Join home team recent form
    left_join(df %>%
                rename_at(vars(matches("[0-9]")),
                          function(x) paste0("home_", str_extract(x, "[a-z][0-9]"))), 
              by = c("date", "season", "home" = "team")) %>%
    # Join away team recent form
    left_join(df %>%
                rename_at(vars(matches("[0-9]")),
                          function(x) paste0("away_", str_extract(x, "[a-z][0-9]"))), 
              by = c("date", "season", "away" = "team")) %>%
    # Remove weeks without at least 1 week sum total - will also remove seasons w/o data
    filter(!is.na(home_r1))
  
  single_lag_results <- list()
  for (i in 1:9) {
    season_results <- list()
    for(this_season in unique(model_df$season)) {
      train <- filter(model_df, season != this_season)
      test <- filter(model_df, season == this_season)
      
      away_reg <- glm(formula(paste0("out ~ home_r", i, 
                                     " + home_t", i,
                                     " + away_r", i, 
                                     " + away_t", i)),
                      data = mutate(train, out = ifelse(res == "A", 1, 0)),
                      family = "binomial")
      draw_reg <- glm(formula(paste0("out ~ home_r", i, 
                                     " + home_t", i,
                                     " + away_r", i, 
                                     " + away_t", i)),
                      data = mutate(train, out = ifelse(res == "D", 1, 0)),
                      family = "binomial")
      home_reg <- glm(formula(paste0("out ~ home_r", i, 
                                     " + home_t", i,
                                     " + away_r", i, 
                                     " + away_t", i)),
                      data = mutate(train, out = ifelse(res == "H", 1, 0)),
                      family = "binomial")
      
      # Predicted output
      season_results[[this_season]] <- train %>%
        mutate(pred_home = predict(home_reg, ., type = "response"),
               pred_away = predict(away_reg, ., type = "response"),
               pred_draw = predict(draw_reg, ., type = "response"),
               log_pred_result = case_when(
                 pred_home > pred_away & pred_home > pred_draw ~ "H",
                 pred_away > pred_home & pred_away > pred_draw ~ "A",
                 pred_draw > pred_home & pred_draw > pred_home ~ "D",
                 TRUE ~ NA_character_
               )) %>%
        filter(!is.na(log_pred_result)) %>%
        summarize(log_train_per = sum(log_pred_result == res) / n()) %>%
        # Test error
        cbind(test %>%
                    mutate(pred_home = predict(home_reg, ., type = "response"),
                           pred_away = predict(away_reg, ., type = "response"),
                           pred_draw = predict(draw_reg, ., type = "response"),
                           log_pred_result = case_when(
                             pred_home > pred_away & pred_home > pred_draw ~ "H",
                             pred_away > pred_home & pred_away > pred_draw ~ "A",
                             pred_draw > pred_home & pred_draw > pred_home ~ "D",
                             TRUE ~ NA_character_
                           )) %>%
                    filter(!is.na(log_pred_result)) %>%
                    summarize(log_test_per = sum(log_pred_result == res) / n()))
      }
    # Take average of all cross-validated measures and save
    single_lag_results[[i]] <- bind_rows(season_results) %>%
      summarize_all(mean)
  }

  # Summarize results
  bind_rows(single_lag_results, .id = "Num Matches") %>%
    select(`Num Matches`, `Logistic Train %` = log_train_per,
           `Logistic Test %` = log_test_per)
}

multi_recent_model <- function(df) {
  # Combine calculated form with observed results
  model_df <- prem_combined %>%
    select(date, season, home, away, res) %>%
    # Join home team recent form
    left_join(df %>%
                rename_at(vars(matches("[0-9]")),
                          function(x) paste0("home_", str_extract(x, "[a-z][0-9]"))), 
              by = c("date", "season", "home" = "team")) %>%
    # Join away team recent form
    left_join(df %>%
                rename_at(vars(matches("[0-9]")),
                          function(x) paste0("away_", str_extract(x, "[a-z][0-9]"))), 
              by = c("date", "season", "away" = "team")) %>%
    # Remove weeks without at least 1 week sum total - will also remove seasons w/o data
    filter(!is.na(home_r1))
  
  multi_lag_results <- list()
  formulas <- vector()
  
  for (i in 1:9) {
    # Set regression formula for models to follow
    formulas[i] <- paste0("+ home_r", i, " + home_t", i,
                               " + away_r", i, " + away_t", i)
    this_reg <- formula(paste0("out ~ ", paste0(formulas[1:i], collapse = " ")))
    
    season_results <- list()
    for(this_season in unique(model_df$season)) {
      train <- filter(model_df, season != this_season)
      test <- filter(model_df, season == this_season)
      
      away_reg <- glm(this_reg,
                      data = mutate(train, out = ifelse(res == "A", 1, 0)),
                      family = "binomial")
      draw_reg <- glm(this_reg,
                      data = mutate(train, out = ifelse(res == "D", 1, 0)),
                      family = "binomial")
      home_reg <- glm(this_reg,
                      data = mutate(train, out = ifelse(res == "H", 1, 0)),
                      family = "binomial")
      
      # Predicted output
      season_results[[this_season]] <- train %>%
        mutate(pred_home = predict(home_reg, ., type = "response"),
               pred_away = predict(away_reg, ., type = "response"),
               pred_draw = predict(draw_reg, ., type = "response"),
               log_pred_result = case_when(
                 pred_home > pred_away & pred_home > pred_draw ~ "H",
                 pred_away > pred_home & pred_away > pred_draw ~ "A",
                 pred_draw > pred_home & pred_draw > pred_home ~ "D",
                 TRUE ~ NA_character_
               )) %>%
        filter(!is.na(log_pred_result)) %>%
        summarize(log_train_per = sum(log_pred_result == res) / n()) %>%
        # Test error
        cbind(test %>%
                    mutate(pred_home = predict(home_reg, ., type = "response"),
                           pred_away = predict(away_reg, ., type = "response"),
                           pred_draw = predict(draw_reg, ., type = "response"),
                           log_pred_result = case_when(
                             pred_home > pred_away & pred_home > pred_draw ~ "H",
                             pred_away > pred_home & pred_away > pred_draw ~ "A",
                             pred_draw > pred_home & pred_draw > pred_home ~ "D",
                             TRUE ~ NA_character_
                           )) %>%
                    filter(!is.na(log_pred_result)) %>%
                    summarize(log_test_per = sum(log_pred_result == res) / n()))
      }
    # Take average of all cross-validated measures and save
    multi_lag_results[[i]] <- bind_rows(season_results) %>%
      summarize_all(mean)
  }
  
  # Summarize results
  bind_rows(multi_lag_results, .id = "Num Matches") %>%
    select(`Num Matches`, `Logistic Train %` = log_train_per,
           `Logistic Test %` = log_test_per)
}
```


```{r read in data and set up summary datasets}
# Read in data
load("Data/prem.Rdata")

prem_combined <- mutate(prem_combined, res = as.factor(res))

# Create dataset of recent goal form for each team throughout season
prem_recent_goals <- prem_combined %>%
  select(date, season, "team" = home, "goals_for" = hg, 
           "goals_against" = ag) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "goals_for" = ag, "goals_against" = hg)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(goalsfor1 = recent_form(goals_for, 1),
         goalsfor2 = recent_form(goals_for, 2),
         goalsfor3 = recent_form(goals_for, 3),
         goalsfor4 = recent_form(goals_for, 4),
         goalsfor5 = recent_form(goals_for, 5),
         goalsfor6 = recent_form(goals_for, 6),
         goalsfor7 = recent_form(goals_for, 7),
         goalsfor8 = recent_form(goals_for, 8),
         goalsfor9 = recent_form(goals_for, 9), 
         goalsagainst1 = recent_form(goals_against, 1),
         goalsagainst2 = recent_form(goals_against, 2),
         goalsagainst3 = recent_form(goals_against, 3),
         goalsagainst4 = recent_form(goals_against, 4),
         goalsagainst5 = recent_form(goals_against, 5),
         goalsagainst6 = recent_form(goals_against, 6),
         goalsagainst7 = recent_form(goals_against, 7),
         goalsagainst8 = recent_form(goals_against, 8),
         goalsagainst9 = recent_form(goals_against, 9)) %>%
  ungroup() %>%
  select(-goals_for, -goals_against)

# Create dataset of recent goal form for each team throughout season
prem_recent_shots <- prem_combined %>%
  filter(!is.na(hs)) %>%
  select(date, season, "team" = home, "shots_for" = hs, 
           "shots_against" = as) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "shots_for" = as, "shots_against" = hs)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(shotsfor1 = recent_form(shots_for, 1),
         shotsfor2 = recent_form(shots_for, 2),
         shotsfor3 = recent_form(shots_for, 3),
         shotsfor4 = recent_form(shots_for, 4),
         shotsfor5 = recent_form(shots_for, 5),
         shotsfor6 = recent_form(shots_for, 6),
         shotsfor7 = recent_form(shots_for, 7),
         shotsfor8 = recent_form(shots_for, 8),
         shotsfor9 = recent_form(shots_for, 9), 
         shotsagainst1 = recent_form(shots_against, 1),
         shotsagainst2 = recent_form(shots_against, 2),
         shotsagainst3 = recent_form(shots_against, 3),
         shotsagainst4 = recent_form(shots_against, 4),
         shotsagainst5 = recent_form(shots_against, 5),
         shotsagainst6 = recent_form(shots_against, 6),
         shotsagainst7 = recent_form(shots_against, 7),
         shotsagainst8 = recent_form(shots_against, 8),
         shotsagainst9 = recent_form(shots_against, 9)) %>%
  ungroup() %>%
  select(-shots_for, -shots_against)

# Create dataset of recent goal form for each team throughout season
prem_recent_shots_target <- prem_combined %>%
  filter(!is.na(hs)) %>%
  select(date, season, "team" = home, "targets_for" = hst, 
           "targets_against" = ast) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "targets_for" = ast, "targets_against" = hst)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(targetsfor1 = recent_form(targets_for, 1),
         targetsfor2 = recent_form(targets_for, 2),
         targetsfor3 = recent_form(targets_for, 3),
         targetsfor4 = recent_form(targets_for, 4),
         targetsfor5 = recent_form(targets_for, 5),
         targetsfor6 = recent_form(targets_for, 6),
         targetsfor7 = recent_form(targets_for, 7),
         targetsfor8 = recent_form(targets_for, 8),
         targetsfor9 = recent_form(targets_for, 9), 
         targetsagainst1 = recent_form(targets_against, 1),
         targetsagainst2 = recent_form(targets_against, 2),
         targetsagainst3 = recent_form(targets_against, 3),
         targetsagainst4 = recent_form(targets_against, 4),
         targetsagainst5 = recent_form(targets_against, 5),
         targetsagainst6 = recent_form(targets_against, 6),
         targetsagainst7 = recent_form(targets_against, 7),
         targetsagainst8 = recent_form(targets_against, 8),
         targetsagainst9 = recent_form(targets_against, 9)) %>%
  ungroup() %>%
  select(-targets_for, -targets_against)

# Create dataset of recent goal form for each team throughout season
prem_recent_corners <- prem_combined %>%
  filter(!is.na(hs)) %>%
  select(date, season, "team" = home, "corners_for" = hc, 
           "corners_against" = ac) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "corners_for" = ac, "corners_against" = hc)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(cornersfor1 = recent_form(corners_for, 1),
         cornersfor2 = recent_form(corners_for, 2),
         cornersfor3 = recent_form(corners_for, 3),
         cornersfor4 = recent_form(corners_for, 4),
         cornersfor5 = recent_form(corners_for, 5),
         cornersfor6 = recent_form(corners_for, 6),
         cornersfor7 = recent_form(corners_for, 7),
         cornersfor8 = recent_form(corners_for, 8),
         cornersfor9 = recent_form(corners_for, 9), 
         cornersagainst1 = recent_form(corners_against, 1),
         cornersagainst2 = recent_form(corners_against, 2),
         cornersagainst3 = recent_form(corners_against, 3),
         cornersagainst4 = recent_form(corners_against, 4),
         cornersagainst5 = recent_form(corners_against, 5),
         cornersagainst6 = recent_form(corners_against, 6),
         cornersagainst7 = recent_form(corners_against, 7),
         cornersagainst8 = recent_form(corners_against, 8),
         cornersagainst9 = recent_form(corners_against, 9)) %>%
  ungroup() %>%
  select(-corners_for, -corners_against)

```


## Various timelines of recent form - single measure
Using logistic regression, look at what timeline of recent goal scoring form best predicts game outcomes using leave-one-season out cross validation and evaluating models on percentage of matches accurately predicted.
```{r recent goals form}

single_recent_model(prem_recent_goals) %>%
  pander(split.cells = 15)

```

Taking into account additional games when determining recent form improves model results, which is unsurprising. Test and training set accuracy are very similar suggesting that the models are not overfitting the data and overall accuracy is moderate, suggesting that adding additional features to the model could result in improved model performance.

## Various timelines of recent form - cumulative measures
Using logistic regression, this section evaluates what timeline of recent form best predicts results, taking into account all shorter measures of recent form. For example, the model of 5-match recent form also includes measures of 3-match and 4-match recent form.
```{r multiple lags same model}
multi_recent_model(prem_recent_goals) %>%
  pander(split.cells = 15)
```


# Recent shots
Using logistic regression, look at what timeline of recent shots attempted best predicts game outcomes using leave-one-season out cross validation and evaluating models on percentage of matches accurately predicted.
```{r recent shots form}
single_recent_model(prem_recent_shots) %>%
  pander(split.cells = 15)
```


## Various timelines of recent shots - cumulative measures
Using logistic regression, this section evaluates what timeline of recent shots best predicts results, taking into account all shorter measures of recent form. For example, the model of 5-match recent shots also includes measures of 3-match and 4-match recent shots.
```{r multiple lags same model shots}
multi_recent_model(prem_recent_shots) %>%
  pander(split.cells = 15)
```


# Recent shots on target
Using logistic regression, look at what timeline of recent shots on target best predicts game outcomes using leave-one-season out cross validation and evaluating models on percentage of matches accurately predicted.
```{r recent shots on target form}
single_recent_model(prem_recent_shots_target) %>%
  pander(split.cells = 15)
```

## Various timelines of recent shots on target - cumulative measures
Using logistic regression, this section evaluates what timeline of recent shots on target best predicts results, taking into account all shorter measures of recent form. For example, the model of 5-match recent shots also includes measures of 3-match and 4-match recent shots.
```{r multiple lags same model shots on target}
multi_recent_model(prem_recent_shots_target) %>%
  pander(split.cells = 15)
```


# Recent corners
Using logistic regression, look at what timeline of recent corners best predicts game outcomes using leave-one-season out cross validation and evaluating models on percentage of matches accurately predicted.
```{r recent corners form}
single_recent_model(prem_recent_corners) %>%
  pander(split.cells = 15)

```

## Various timelines of recent corners - cumulative measures
Using logistic regression, this section evaluates what timeline of recent shots on target best predicts results, taking into account all shorter measures of recent form. For example, the model of 5-match recent shots also includes measures of 3-match and 4-match recent shots.
```{r multiple lags same model corners}
multi_recent_model(prem_recent_corners) %>%
  pander(split.cells = 15)
```