---
title: "Premier League Multiple Features Match Result"
author: "Craig McGowan"
date: "June 29, 2018"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(pander)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "../")
```

```{r functions}
# Recent form for a particular stat
recent_form <- function(var, num_weeks) {

  lag(zoo::rollsum(var, num_weeks, na.pad = TRUE, align = "right"), 1)
 
}

single_recent_model <- function(df, var_order) {
  
  model_df <- prem_combined %>%
    select(date, season, home, away, res) %>%
    # Join home team recent form
    left_join(df %>%
                rename_at(vars(matches("[0-9]")),
                          function(x) paste0("home_", x)), 
              by = c("date", "season", "home" = "team")) %>%
    # Join away team recent form
    left_join(df %>%
                rename_at(vars(matches("[0-9]")),
                          function(x) paste0("away_", x)), 
              by = c("date", "season", "away" = "team")) %>%
    # Remove weeks without prediction data
    filter(!is.na(home_targetsfor4))
  
  model_results <- list()
  for (i in 1:length(var_order)) {
    season_results <- list()
    
    # Build regression equation
    reg_string <- "out ~ "
    for(j in 1:i) {
      reg_string <- paste0(reg_string, "+ home_", var_order[j], "for4 + home_", 
                           var_order[j], "against4 + away_", var_order[j],
                           "for4 + away_", var_order[j], "against4")
    }
    
    for(this_season in unique(model_df$season)) {
      train <- filter(model_df, season != this_season)
      test <- filter(model_df, season == this_season)
      
      away_reg <- glm(formula(reg_string),
                      data = mutate(train, out = ifelse(res == "A", 1, 0)),
                      family = "binomial")
      draw_reg <- glm(formula(reg_string),
                      data = mutate(train, out = ifelse(res == "D", 1, 0)),
                      family = "binomial")
      home_reg <- glm(formula(reg_string),
                      data = mutate(train, out = ifelse(res == "H", 1, 0)),
                      family = "binomial")
      
      # Predicted output
      season_results[[this_season]] <- train %>%
        mutate(pred_home = predict(home_reg, ., type = "response"),
               pred_away = predict(away_reg, ., type = "response"),
               pred_draw = predict(draw_reg, ., type = "response"),
               log_pred_result = case_when(
                 pred_home > pred_away & pred_home > pred_draw ~ "H",
                 pred_away > pred_home & pred_away > pred_draw ~ "A",
                 pred_draw > pred_home & pred_draw > pred_home ~ "D",
                 TRUE ~ NA_character_
               )) %>%
        filter(!is.na(log_pred_result)) %>%
        summarize(log_train_per = sum(log_pred_result == res) / n()) %>%
        # Test error
        cbind(test %>%
                    mutate(pred_home = predict(home_reg, ., type = "response"),
                           pred_away = predict(away_reg, ., type = "response"),
                           pred_draw = predict(draw_reg, ., type = "response"),
                           log_pred_result = case_when(
                             pred_home > pred_away & pred_home > pred_draw ~ "H",
                             pred_away > pred_home & pred_away > pred_draw ~ "A",
                             pred_draw > pred_home & pred_draw > pred_home ~ "D",
                             TRUE ~ NA_character_
                           )) %>%
                    filter(!is.na(log_pred_result)) %>%
                    summarize(log_test_per = sum(log_pred_result == res) / n()))
      }
    # Take average of all cross-validated measures and save
    model_results[[var_order[i]]] <- bind_rows(season_results) %>%
      summarize_all(mean)
  }

  # Summarize results
  bind_rows(model_results, .id = "Last variable added") %>%
    select(`Last variable added`, `Logistic Train %` = log_train_per,
           `Logistic Test %` = log_test_per)
}

```

```{r read in data and set up summary datasets}
# Read in data
load("Data/prem.Rdata")

prem_combined <- mutate(prem_combined, res = as.factor(res))

# Create dataset of recent goal form for each team throughout season
prem_recent_goals <- prem_combined %>%
  select(date, season, "team" = home, "goals_for" = hg, 
           "goals_against" = ag) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "goals_for" = ag, "goals_against" = hg)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(goalsfor4 = recent_form(goals_for, 4),
         goalsagainst4 = recent_form(goals_against, 4)) %>%
  ungroup() %>%
  select(-goals_for, -goals_against)

# Create dataset of recent goal form for each team throughout season
prem_recent_shots <- prem_combined %>%
  filter(!is.na(hs)) %>%
  select(date, season, "team" = home, "shots_for" = hs, 
           "shots_against" = as) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "shots_for" = as, "shots_against" = hs)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(shotsfor4 = recent_form(shots_for, 4),
         shotsagainst4 = recent_form(shots_against, 4)) %>%
  ungroup() %>%
  select(-shots_for, -shots_against)

# Create dataset of recent goal form for each team throughout season
prem_recent_shots_target <- prem_combined %>%
  filter(!is.na(hst)) %>%
  select(date, season, "team" = home, "targets_for" = hst, 
           "targets_against" = ast) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "targets_for" = ast, "targets_against" = hst)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(targetsfor4 = recent_form(targets_for, 4),
         targetsagainst4 = recent_form(targets_against, 4)) %>%
  ungroup() %>%
  select(-targets_for, -targets_against)

# Create dataset of recent goal form for each team throughout season
prem_recent_corners <- prem_combined %>%
  filter(!is.na(hc)) %>%
  select(date, season, "team" = home, "corners_for" = hc, 
           "corners_against" = ac) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "corners_for" = ac, "corners_against" = hc)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(cornersfor4 = recent_form(corners_for, 4),
         cornersagainst4 = recent_form(corners_against, 4)) %>%
  ungroup() %>%
  select(-corners_for, -corners_against)

# Create dataset of recent fouls for each team throughout season
prem_recent_fouls <- prem_combined %>%
  filter(!is.na(hf)) %>%
  select(date, season, "team" = home, "fouls_for" = hf, 
           "fouls_against" = af) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "fouls_for" = af, "fouls_against" = hf)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(foulsfor4 = recent_form(fouls_for, 4),
         foulsagainst4 = recent_form(fouls_against, 4)) %>%
  ungroup() %>%
  select(-fouls_for, -fouls_against)

# # Create dataset of recent offsides for each team throughout season
# prem_recent_offsides <- prem_combined %>%
#   filter(!is.na(ho)) %>%
#   select(date, season, "team" = home, "offsides_for" = ho, 
#            "offsides_against" = ao) %>%
#     bind_rows(prem_combined %>%
#                 select(date, season, "team" = away,
#                        "offsides_for" = ao, "offsides_against" = ho)) %>%
#   group_by(season, team) %>%
#   arrange(date, .by_group = TRUE) %>%
#   mutate(offsidesfor4 = recent_form(offsides_for, 4),
#          offsidesagainst4 = recent_form(offsides_against, 4)) %>%
#   ungroup() %>%
#   select(-offsides_for, -offsides_against)

# Create dataset of recent offsides for each team throughout season
prem_recent_ycs <- prem_combined %>%
  filter(!is.na(hy)) %>%
  select(date, season, "team" = home, "yc_for" = hy, 
           "yc_against" = ay) %>%
    bind_rows(prem_combined %>%
                select(date, season, "team" = away,
                       "yc_for" = ay, "yc_against" = hy)) %>%
  group_by(season, team) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(ycfor4 = recent_form(yc_for, 4),
         ycagainst4 = recent_form(yc_against, 4)) %>%
  ungroup() %>%
  select(-yc_for, -yc_against)

# Create dataset combining all recent measures together
prem_recent_all <- prem_recent_goals %>%
  inner_join(prem_recent_shots, by = c("date", "season", "team")) %>%
  inner_join(prem_recent_shots_target, by = c("date", "season", "team")) %>%
  inner_join(prem_recent_corners, by = c("date", "season", "team")) %>%
  inner_join(prem_recent_fouls, by = c("date", "season", "team")) %>%
  # inner_join(prem_recent_offsides, by = c("date", "season", "team")) %>%
  inner_join(prem_recent_ycs, by = c("date", "season", "team"))

```

# Combine goals scored, shots, and shots on target
Individual components have maxed out at around 52% of games being predicted correctly. Next, we combine multiple measures looked at so far (goals scored, shots, and shots on target). Using logistic regression, look at what timeline of recent form as measured by all metrics best predicts game outcomes using leave-one-season out cross validation and evaluating models on percentage of matches accurately predicted.
```{r recent goals shots and shots on target}
# Combine calculated form with observed results
check_order <- c("shots", "targets", "corners", "goals", "fouls", "yc")

mult_feature_result <- single_recent_model(prem_recent_all, check_order)

pander(mult_feature_result)

gather(mult_feature_result, key = "train_test", value = "perform", `Logistic Train %`, 
         `Logistic Test %`) %>%
  mutate(`Last variable added` = factor(`Last variable added`,
                                        levels = check_order)) %>%
  ggplot(aes(x = `Last variable added`, y = perform)) +
  geom_point(aes(shape = train_test))

```